### 2023-11-20 script.qmd
scripts file and data form

---
title: "script"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-11-21 scripts/manuscript.qmd
codebook for selected important variables and gss2022 data

---
title: "Social Conditions of Fear"
author:
  - Alejandra Pedraza
  - W. Jamie Yang
format: html
highlight-style: pygments
editor: visual
execute:
  eval: true # change to false before final submission
  echo: true
  message: false
  warning: false
code-fold: true
embed-resources: true
bibliography: references.bib
---

# codebook for variables

**Dependent Variable**
"fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

**Sociodemographics Variables**
1. "born": Were you born in this country?
2. "age"
3. "sexnow1": Do you describe yourself as male, female, or transgender? 
4. "sexornt": Which of the following best describes you? 
5. "racerank1": If you had to choose, which of these races do you identify most with?
6. "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
7. "degree": Respondent's degree.
8. "uscitzn": Are you a US citizen?
9. "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?

**Other Controls**
1 "attend": How often do you attend religious services?
2. "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
3. "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
4. "health": Would you say your own health, in general, is excellent, good, fair, or poor?
5. "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
6. "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
7. "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(dplyr)
library(tidyr)
library(here)
library(rio)
library(janitor)
library(knitr)
library(kableExtra)
```

```{r}
#| label: loading data

gss <- import(here("data", "GSS.xlsx")) %>%
  as_tibble() %>% 
  characterize() %>% 
  clean_names()

view(gss)
```

For the purposes of EDLD 651 final project, we only use data from the year 2022. This will result in a much smaller data set, and we won't need to work with panel models for now.

```{r}
#| label: gss 2022
gss2022 <- gss %>% 
  filter(year == "2022")

view(gss2022)
```

I noticed that there are 3544 entries, but the  variable "id" goes to 3545. After investigation, I found that "id == 3419" doesn't exist and jumps from 3418 to 3420. It doesn't affect our work at the moment, so I'll leave it for now (and fix it later when needed). "id" is a functional unique identifier in our data.

### 2023-11-21 scripts/manuscript.qmd
select the variables we need

@@
 3. "sexnow1": Do you describe yourself as male, female, or transgender? 
+*This is self-described gender.*
+4. "sex": Code respondent's sex
+*This is gender presentation as interpreted by the interviewer.*
 4. "sexornt": Which of the following best describes you? 
 5. "racerank1": If you had to choose, which of these races do you identify most with? 
+*Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
 6. "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
@@
 7. "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+8. "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+
+# set up
 
@@
 gss2022 <- gss %>% 
-  filter(year == "2022")
+  filter(year == "2022") %>% 
+  select("year", "id", "age", "sex", "sexnow1", "sex", "degree", "race", "born", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "sexornt", "hispanic", "racerank1", "uscitzn", "dwelown")
 
@@
 "id" is a functional unique identifier in our data.
+
+# recode variables

### 2023-11-21 scripts/manuscript.qmd
select needed variables into analytical data set

--- 2023-11-21 scripts/manuscript.qmd
@@
 "id" is a functional unique identifier in our data.
+
+# analytical dataset
+
+```{r}
+#| label: select variables
+data <- gss2022 %>% 
+    select("year", "id", "age", "sex", "sexnow1", "sex", "degree", "race", "born", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "sexornt", "hispanic", "racerank1", "uscitzn", "dwelown")
+
+view(data)
+```
+
+```{r}
+#| label: recode variables
+
+```

### 2023-11-21 scripts/manuscript.qmd
racecombo variable, deleted na's from dependent variable and also race variable

@@
 
-**Dependent Variable**
-"fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+**Dependent Variable** "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
-**Sociodemographics Variables**
-1. "born": Were you born in this country?
-2. "age"
-3. "sexnow1": Do you describe yourself as male, female, or transgender? 
-*This is self-described gender.*
-4. "sex": Code respondent's sex
-*This is gender presentation as interpreted by the interviewer.*
-4. "sexornt": Which of the following best describes you? 
-5. "racerank1": If you had to choose, which of these races do you identify most with? 
-*Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
-6. "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
-7. "degree": Respondent's degree.
-8. "uscitzn": Are you a US citizen?
-9. "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
+**Sociodemographics Variables** 1.
+"born": Were you born in this country?
+2.
+"age" 3.
+"sexnow1": Do you describe yourself as male, female, or transgender?
+*This is self-described gender.* 4.
+"sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.* 4.
+"sexornt": Which of the following best describes you?
+5.
+"racerank1": If you had to choose, which of these races do you identify most with?
+*Kept this race variable here instead of "race" because it allows more heterogeneity within group.* 6.
+"hispanic": Are you Spanish, Hispanic, or Latino/Latina?
+IF YES: Which group are you from?
+7.
+"degree": Respondent's degree.
+8.
+"uscitzn": Are you a US citizen?
+9.
+"dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
-**Other Controls**
-1 "attend": How often do you attend religious services?
-2. "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
-3. "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
-4. "health": Would you say your own health, in general, is excellent, good, fair, or poor?
-5. "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
-6. "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
-7. "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
-8. "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+**Other Controls** 1 "attend": How often do you attend religious services?
+2.
+"raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+3.
+"happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+4.
+"health": Would you say your own health, in general, is excellent, good, fair, or poor?
+5.
+"helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+6.
+"fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+7.
+"trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+8.
+"partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
@@
 
+Below, we create the analytical dataset.
+We use the gss2022 subset we created above, leaving the original dataset untouched -- this is important for replication purposes.
+First, we dropped the missing values for our dependent variable.
+The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
+For the purposes of this project, we are considering all of the above as missing values.
+
+Then we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This was not going to work.
+So, we opted to use the "racerank1" variable, since in this case participants were able to answer with either a race or ethnicity.
+There were 16 different responses: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
+We recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
+We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
+Additionally, because race is an important control that we identify, we dropped the missing values.
+
 ```{r}
@@
 data <- gss2022 %>% 
-    select("year", "id", "age", "sex", "sexnow1", "sex", "degree", "race", "born", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "sexornt", "hispanic", "racerank1", "uscitzn", "dwelown")
+    select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racerank1", "uscitzn", "dwelown")
 
@@
 #| label: recode variables
+tidydata <- data %>% 
+  mutate(fear = 
+           case_when( fear == ".i:  Inapplicable" | fear == ".n: No answer" | 
+                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ "missing", 
+                      fear == "YES" ~ "yes", 
+                      fear == "NO" ~ "no")) %>% 
+  filter(fear == "yes" | fear == "no") %>% 
+  mutate(racecombo = 
+           case_when(racerank1 == "White" ~ "White",
+                     racerank1 == "Black or African American" ~ "Black",
+                     racerank1 == "Hispanic" ~ "Hispanic",
+                     racerank1 == "Asian Indian" | racerank1 ==  "Chinese" |
+                       racerank1 ==  "Filipino" | racerank1 ==  "Japanese" | 
+                       racerank1 ==  "Korean" |  racerank1 ==  "Vietnamese" |
+                       racerank1 ==  "Other Asian" ~ "Asian",
+                     racerank1 == "American Indian or Alaska Native" ~ "AIAN",
+                     racerank1 == "Guamanian or Chamorro" | racerank1 == "Samoan" |
+                       racerank1 == "Other Pacific Islander" ~ "NHPI",
+                      racerank1 == "Some other race" ~ "Other",
+                     .default = "missing")) %>% 
+  filter(racecombo != "missing") %>% 
+  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")
 

### 2023-11-22 scripts/manuscript.qmd
attempting to deal with missing values

@@
 library(kableExtra)
+library(mice)
 ```
@@
   as_tibble() %>% 
-  characterize() %>% 
+  characterize() %>% # might not need characterize, TBD
   clean_names()
@@
 ```{r}
-#| label: select variables
+#| label: select-variables
 data <- gss2022 %>% 
-    select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racerank1", "uscitzn", "dwelown")
+    select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
 
@@
 ```{r}
-#| label: recode variables
+#| label: coding all missing data as NA 
+# NOT WORKING - Need to troubleshoot
+missing_vec <- c(".i:  Inapplicable" , ".n: No answer", ".d: Do not Know/Cannot Choose", ".s: Skipped on Web")
+
+no_na_data <- data %>% 
+  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), ~ifelse(.
+%in% missing_vec, NA_character_, .)))
+
+#data_complete <- mice(data, m=5)
+
+#summary(complete(data,1))
+
+```
+
+```{r}
+#| label: mice imputations to deal with missing data
+
+#data_complete <- mice(no_na_data, m=5)
+
+#summary(complete(data_complete,1))
+```
+
+```{r}
+#| label: recode-variables
 tidydata <- data %>% 
@@
            case_when( fear == ".i:  Inapplicable" | fear == ".n: No answer" | 
-                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ "missing", 
+                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ NA_character_, 
                       fear == "YES" ~ "yes", 
@@
                       racerank1 == "Some other race" ~ "Other",
-                     .default = "missing")) %>% 
-  filter(racecombo != "missing") %>% 
+                     .default = NA_character_)) %>% 
+  filter(!is.na(racecombo)) %>% 
   select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")

### 2023-11-22 scripts/manuscript.qmd
cleaning

@@
 # NOT WORKING - Need to troubleshoot
-missing_vec <- c(".i:  Inapplicable" , ".n: No answer", ".d: Do not Know/Cannot Choose", ".s: Skipped on Web")
+missing_vec <- c(".i:  Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", ".s: Skipped on Web")
 
-no_na_data <- data %>% 
+na_data <- data %>% 
   mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), ~ifelse(.

### 2023-11-22 scripts/manuscript.qmd
Replacing NA values with "NA" instead of different wordings as originally in the data

@@
 
-**Dependent Variable** "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+**Dependent Variable** 
+"fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
-**Sociodemographics Variables** 1.
-"born": Were you born in this country?
-2.
-"age" 3.
-"sexnow1": Do you describe yourself as male, female, or transgender?
-*This is self-described gender.* 4.
-"sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.* 4.
-"sexornt": Which of the following best describes you?
-5.
-"racerank1": If you had to choose, which of these races do you identify most with?
-*Kept this race variable here instead of "race" because it allows more heterogeneity within group.* 6.
-"hispanic": Are you Spanish, Hispanic, or Latino/Latina?
-IF YES: Which group are you from?
-7.
-"degree": Respondent's degree.
-8.
-"uscitzn": Are you a US citizen?
-9.
-"dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
+**Sociodemographics Variables** 
+1. "born": Were you born in this country? 
+2. "age" 
+3. "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.* 
+4. "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.* 
+5. "sexornt": Which of the following best describes you? 
+6. "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.* 
+7. "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from? 
+8. "degree": Respondent's degree. 
+9. "uscitzn": Are you a US citizen? 
+10. "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
-**Other Controls** 1 "attend": How often do you attend religious services?
-2.
-"raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
-3.
-"happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
-4.
-"health": Would you say your own health, in general, is excellent, good, fair, or poor?
-5.
-"helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
-6.
-"fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
-7.
-"trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
-8.
-"partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+**Other Controls** 
+1 "attend": How often do you attend religious services? 
+2. "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now? 
+3. "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy? 
+4. "health": Would you say your own health, in general, is excellent, good, fair, or poor? 
+5. "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves? 
+6. "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair? 
+7. "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people? 
+8. "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
@@
 na_data <- data %>% 
-  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), ~ifelse(.
-%in% missing_vec, NA_character_, .)))
+  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
+                ~ifelse(. %in% missing_vec, NA_character_, .)))
+
+view(na_data)
 

### 2023-11-22 scripts/manuscript.qmd
mice imputation

@@
 #| label: coding all missing data as NA
-# NOT WORKING - Need to troubleshoot
+
 missing_vec <- c(".i:  Inapplicable", ".n:  No answer", ".d:  Do not Know/Cannot Choose", ".s:  Skipped on Web")
@@
 
-#data_complete <- mice(data, m=5)
+```
 
-#summary(complete(data,1))
+```{r}
+#| label: mice imputation
+
+complete5 <- mice(na_data, 
+                 m=5, 
+                 seed = 25)
+
+complete <- complete(complete5,1)
 
@@
 ```{r}
-#| label: mice imputations to deal with missing data
-
-#data_complete <- mice(no_na_data, m=5)
-
-#summary(complete(data_complete,1))
-```
-
-```{r}
 #| label: recode-variables
-tidydata <- data %>% 
-  mutate(fear = 
-           case_when( fear == ".i:  Inapplicable" | fear == ".n: No answer" | 
-                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ NA_character_, 
-                      fear == "YES" ~ "yes", 
-                      fear == "NO" ~ "no")) %>% 
-  filter(fear == "yes" | fear == "no") %>% 
+tidydata <- na_data %>% 
+  filter(!is.na(fear)) %>% 
   mutate(racecombo = 
@@
                        racerank1 == "Other Pacific Islander" ~ "NHPI",
-                      racerank1 == "Some other race" ~ "Other",
-                     .default = NA_character_)) %>% 
-  filter(!is.na(racecombo)) %>% 
+                      racerank1 == "Some other race" ~ "Other")) %>% 
   select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")

### 2023-11-22 scripts/manuscript.qmd
mice tries
reformatted document

@@
   - W. Jamie Yang
+abstract: "TBW" # this is where abstract should be pasted 
 format: html
@@
 code-fold: true
+code-summary: "Show the code"
 embed-resources: true
@@
 
-# codebook for variables
+# Introduction
+
+# Methods
+
+# Results
+
+# Discussion
+
+# Codebook
 
 **Dependent Variable**
-"fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+-   "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
 **Sociodemographics Variables**
+
 1.  "born": Were you born in this country?
+
 2.  "age"
+
 3.  "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+
 4.  "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
+
 5.  "sexornt": Which of the following best describes you?
+
 6.  "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
+
 7.  "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
+
 8.  "degree": Respondent's degree.
+
 9.  "uscitzn": Are you a US citizen?
+
 10. "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
@@
 **Other Controls**
-1 "attend": How often do you attend religious services? 
+
+1.  "attend": How often do you attend religious services?
+
 2.  "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+
 3.  "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+
 4.  "health": Would you say your own health, in general, is excellent, good, fair, or poor?
+
 5.  "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+
 6.  "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+
 7.  "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+
 8.  "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
-# set up
+# Data Cleaning
 
@@
 library(mice)
+library(forcats)
 ```
@@
 #| label: loading data 
-
 gss <- import(here("data", "GSS.xlsx")) %>%
@@
 
-For the purposes of EDLD 651 final project, we only use data from the year 2022. This will result in a much smaller data set, and we won't need to work with panel models for now.
-
 ```{r}
-#| label: gss 2022
+#| label: filtering 2022 data
 gss2022 <- gss %>% 
@@
 ```
-
-I noticed that there are 3544 entries, but the variable "id" goes to 3545.
-After investigation, I found that "id == 3419" doesn't exist and jumps from 3418 to 3420.
-It doesn't affect our work at the moment, so I'll leave it for now (and fix it later when needed).
-"id" is a functional unique identifier in our data.
-
-# analytical dataset
-
-Below, we create the analytical dataset.
-We use the gss2022 subset we created above, leaving the original dataset untouched -- this is important for replication purposes.
-First, we dropped the missing values for our dependent variable.
-The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
-For the purposes of this project, we are considering all of the above as missing values.
-
-Then we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This was not going to work.
-So, we opted to use the "racerank1" variable, since in this case participants were able to answer with either a race or ethnicity.
-There were 16 different responses: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
-We recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
-We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
-Additionally, because race is an important control that we identify, we dropped the missing values.
 
@@
 ```
+
+The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
+For the purposes of this project, we are considering all of the above as missing values.
+Below we recode all these indicators as NAs.
+This is necessary to rice mice imputations to address missing values.
 
@@
 
+To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
+
 ```{r}
@@
 complete5 <- mice(na_data, 
-                 m=5, 
+                 m = 50, 
                  seed = 25)
 
-complete <- complete(complete5,1)
+summary(complete(complete5, 1))
+
+temp <- complete(complete5, action = "long", include = TRUE)
+
+tidydata_complete <- as.mids(temp)
+
+save(tidydata_complete, file = "data/tidydata_complete.RData")
 
@@
 
+Race/ ethnicity is an important control for our models.
+First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
+So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
+We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
+We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
+
 ```{r}
-#| label: recode-variables
+#| label: recode variables
 tidydata <- na_data %>% 
@@
   mutate(racecombo = 
-           case_when(racerank1 == "White" ~ "White",
-                     racerank1 == "Black or African American" ~ "Black",
+           factor(case_when(racerank1 == "White" ~ "White",
+                     racerank1 == "Black or African-American" ~ "Black",
                      racerank1 == "Hispanic" ~ "Hispanic",
@@
                        racerank1 == "Other Pacific Islander" ~ "NHPI",
-                      racerank1 == "Some other race" ~ "Other")) %>% 
-  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")
+                      racerank1 == "Some other race" ~ "Other"))) %>% 
+  mutate(racecombo = fct_relevel(racecombo, 
+              "White",
+              "Black", 
+              "Asian",
+              "Hispanic",
+              "AIAN",
+              "NHPI",
+              "Other")) %>% 
+  select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
 

### 2023-11-28 scripts/manuscript.qmd

@@
 
+#XM: I understand this is a draft version and would suggest adding more information about the data set.
+
 **Dependent Variable**
@@
 3.  "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+
+    #XM: To me this question really doesn't make sense... there are transgender males and females, did they mean non-binary? I'd feel concerned about this variable
 
@@
 
+#XM: there are still NAs in the imputated data even after I tried changing parameters in mice(), maybe reducing the variables would help?
+also I think it might be helpful to investigate missing patterns and correlations before imputations
 summary(complete(complete5, 1))

### 2023-11-28 scripts/manuscript.qmd
recoding variables for analysis

--- 2023-11-22 scripts/manuscript.qmd
@@
 
-1.  "born": Were you born in this country?
+-   "born": Were you born in this country?
 
-2.  "age"
+-   "age"
 
-3.  "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+-   "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
 
-4.  "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
+-   "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
 
-5.  "sexornt": Which of the following best describes you?
+-   "sexornt": Which of the following best describes you?
 
-6.  "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
+-   "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
 
-7.  "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
+-   "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
 
-8.  "degree": Respondent's degree.
+-   "degree": Respondent's degree.
 
-9.  "uscitzn": Are you a US citizen?
+-   "uscitzn": Are you a US citizen?
 
-10. "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
+-   "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
@@
 
-1.  "attend": How often do you attend religious services?
+-   "attend": How often do you attend religious services?
 
-2.  "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+-   "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
 
-3.  "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+-   "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
 
-4.  "health": Would you say your own health, in general, is excellent, good, fair, or poor?
+-   "health": Would you say your own health, in general, is excellent, good, fair, or poor?
 
-5.  "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+-   "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
 
-6.  "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+-   "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
 
-7.  "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+-   "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
 
-8.  "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+-   "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
@@
   as_tibble() %>% 
-  characterize() %>% # might not need characterize, TBD
   clean_names()
 
-view(gss)
 ```
@@
 
-view(gss2022)
 ```
@@
 data <- gss2022 %>% 
-    select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
+    select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
 
-view(data)
 ```
@@
 Below we recode all these indicators as NAs.
-This is necessary to rice mice imputations to address missing values.
+This is necessary to run mice imputations to address missing values.
 
@@
 na_data <- data %>% 
-  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
+  mutate(across(c(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
                 ~ifelse(. %in% missing_vec, NA_character_, .)))
-
-view(na_data)
 
@@
 #| label: mice imputation
-
+# DO NOT RUN 
 complete5 <- mice(na_data, 
@@
 
+Fear is our dependent variable, so the first step was to remove NAs for that variabble.
+
 Race/ ethnicity is an important control for our models.
@@
 ```{r}
-#| label: recode variables
+#| label: recoding variables 
 tidydata <- na_data %>% 
@@
               "Other")) %>%   
-  select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
+  mutate(partnrs5 = 
+           factor(case_when(partnrs5 == "NO PARTNERS" ~ "0",
+                            partnrs5 == "1 PARTNER" | partnrs5 == "2 PARTNERS" |
+                              partnrs5 == "3 PARTNERS" | 
+                              partnrs5 == "4 PARTNERS" ~ "Less than 5",
+                            partnrs5 == "5-10 PARTNERS" ~ "5-10",
+                            partnrs5 == "11-20 PARTNERS" ~ "11-20",
+                            partnrs5 == "21-100 PARTNERS" ~ "21-100",
+                            partnrs5 == "MORE THAN 100 PARTNERS" ~ "100 or more",
+                            .default = NA))) %>%
+  mutate(partnrs5 = fct_relevel(partnrs5,
+                                "0",
+                                "Less than 5",
+                                "5-10",
+                                "11-20",
+                                "21-100",
+                                "100 or more")) %>%
+  mutate(uscitzn = 
+           factor(case_when(uscitzn == "A u.s. citizen" | uscitzn == "A u.s. citizen born in puerto rico, the u.s. virgin islands, or the northern marianas islands (if volunteered)" |
+                       uscitzn == "Born outside of the u.s. to parents who were u.s. citizens at that time (if volunteered)" ~ "Yes",
+                     uscitzn == "Not a u.s. citizen" ~ "No",
+                     .default = NA))) %>% 
+   mutate(dwelown =
+           factor(case_when(dwelown == "Own or is buying" ~ "Owner",
+                            dwelown == "Pays rent" ~ "Renter",
+                            dwelown == "Other" ~ "Other"))) %>% 
+  mutate(degree = fct_relevel(degree, 
+                              "Less than high school",
+                              "High school",
+                              "Associate/junior college",
+                              "Bachelor's",
+                              "Graduate")) %>% 
+  mutate(attend = fct_relevel(attend, 
+                              "Never",
+                              "Less than once a year",
+                              "About once or twice a year",
+                              "2-3 times a month",
+                              "Nearly every week",
+                              "Every week",
+                              "Several times a week")) %>% 
+  mutate(happy = fct_relevel(happy, 
+                             "Not too happy",
+                             "Pretty happy",
+                             "Very happy",)) %>% 
+  mutate(health = fct_relevel(health, 
+                              "Poor",
+                              "Fair",
+                              "Good",
+                              "Excellent")) %>% 
+  mutate(helpful = fct_relevel(helpful, 
+                               "Looking out for themselves",
+                               "Try to be helpful",
+                               "Depends")) %>% 
+  mutate(fair = fct_relevel(fair, 
+                            "Would take advantage of you",
+                            "Would try to be fair",
+                            "Depends")) %>% 
+  mutate(trust = fct_relevel(trust, 
+                             "Can't be too careful",
+                             "Can trust",
+                             "Depends")) %>% 
+  select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
 
 ```
+
+```{r}
+```

### 2023-11-28 scripts/manuscript.qmd
figure 1

@@
 library(forcats)
+library(ggplot2)
 ```
@@
 #| label: mice imputation
-# DO NOT RUN 
+#| eval: false
 complete5 <- mice(na_data, 
@@
   filter(!is.na(fear)) %>% 
+  mutate(fear = factor(case_when(fear == "NO" ~ "no",
+                                 fear == "YES" ~ "yes"))) %>% 
   mutate(racecombo = 
@@
 ```{r}
+#| label: creating a smaller subset from data
+
+gss_subset <- tidydata %>% 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
+         partnrs5, fear) 
 ```
+
+```{r}
+#| label: pivot longer and pivot wider to get proportion of yes and no by race and display on a table
+#| eval: false
+
+temp <- gss_subset %>% 
+  pivot_longer(
+    cols = fear,
+    names_to = "fear",
+    values_to = "answer") %>% 
+  select(racecombo, answer) %>% 
+  mutate(fear = as.numeric(case_when(answer == "no" ~ "0",
+                                 answer == "yes" ~ "1"))) %>% 
+  select(racecombo, fear)
+
+%>% 
+  pivot_wider(
+    names_from = racecombo,
+    values_from = fear
+  )
+
+%>% 
+  group_by(racecombo)
+
+%>% 
+  summarize(proportion = fear /sum(fear == "yes", na.rm = TRUE)) %>% 
+  kable(caption = "Proportion Social Fear by Race",
+        col.names = c("Race", "Proportion"),
+        digits = 1,
+        booktabs = TRUE)
+
+```
+
+```{r}
+#| label: fig-fear_race
+#| fig-cap: Bar graph showing the proportion or respondants, by race, that feared walking alone at night.
+
+tidydata %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
+             fill = fear)) +
+  geom_bar(position = "dodge") + 
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Race",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Race",
+       y = NULL,
+       fill = "Fear"
+       ) +
+  theme_bw()
+
+```

### 2023-11-28 scripts/manuscript.qmd
table

--- 2023-11-28 scripts/manuscript.qmd
@@
 #| label: mice imputation
-# DO NOT RUN 
+#| eval: false
 complete5 <- mice(na_data, 
-                 m = 50, 
+                 m = 5, 
                  seed = 25)
@@
 ```{r}
+#| label: creating a smaller subset from data
+
+gss_subset <- tidydata %>% 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
+         partnrs5, fear) 
 ```
+
+```{r}
+#| label: tbl-fear_race
+#| tbl-cap: Table showing mean fear by race.
+
+gss_subset %>% 
+  pivot_longer(
+    cols = fear,
+    names_to = "fear",
+    values_to = "answer") %>% 
+  select(racecombo, answer) %>% 
+  mutate(fear = as.numeric(case_when(answer == "no" ~ "0",
+                                 answer == "yes" ~ "1"))) %>% 
+  select(racecombo, fear) %>% 
+  group_by(racecombo) %>% 
+  summarize(mean_fear = mean(fear)) %>% 
+  kable(caption = "Mean Social Fear by Race",
+        col.names = c("Race", "Mean"),
+        digits = 3)
+```
+
+```{r}
+#| label: fig-fear_race
+#| fig-cap: Bar graph showing the proportion or respondants, by race, that feared walking alone at night.
+
+gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
+             fill = fear)) +
+  geom_bar(position = "dodge") + 
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Race",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Race",
+       y = NULL,
+       fill = "Fear"
+       ) +
+  theme_bw()
+
+```

### 2023-11-29 scripts/manuscript.qmd
last commit
adjusted formatting of socio-demographic variables to numeric list, suggested the following: using head() instead of view(), using fct_recode instead of code_when to recode race categories, using skim inspect data.

--- 2023-11-28 scripts/manuscript.qmd
@@
   - W. Jamie Yang
-abstract: "TBW" # this is where abstract should be pasted 
 format: html
@@
 code-fold: true
-code-summary: "Show the code"
 embed-resources: true
@@
 
-# Introduction
+# codebook for variables
 
-# Methods
-
-# Results
-
-# Discussion
-
-# Codebook
-
-**Dependent Variable**
-
--   "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+**Dependent Variable** "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
@@
 
--   "born": Were you born in this country?
+1.  "born": Were you born in this country?
 
--   "age"
+2.  "age"
 
--   "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+3.  "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
 
--   "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
+4.  "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
 
--   "sexornt": Which of the following best describes you?
+5.  "sexornt": Which of the following best describes you?
 
--   "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
+6.  "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
 
--   "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
+7.  "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from? 8. "degree": Respondent's degree.
 
--   "degree": Respondent's degree.
+8.  "uscitzn": Are you a US citizen?
 
--   "uscitzn": Are you a US citizen?
+9.  "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
--   "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
+# I modified the formatting of this list so that it appears numerically.
 
@@
 
--   "attend": How often do you attend religious services?
+1.  "attend": How often do you attend religious services?
 
--   "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+2.  "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
 
--   "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+3.  "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
 
--   "health": Would you say your own health, in general, is excellent, good, fair, or poor?
+4.  "health": Would you say your own health, in general, is excellent, good, fair, or poor?
 
--   "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+5.  "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
 
--   "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+6.  "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
 
--   "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+7.  "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
 
--   "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+8.  "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
-# Data Cleaning
+# set up
 
@@
 library(mice)
-library(forcats)
 ```
@@
 #| label: loading data
+
 gss <- import(here("data", "GSS.xlsx")) %>%
   as_tibble() %>% 
+  characterize() %>% # might not need characterize, TBD
   clean_names()
 
+view(gss)
 ```
 
+For the purposes of EDLD 651 final project, we only use data from the year 2022. This will result in a much smaller data set, and we won't need to work with panel models for now.
+
 ```{r}
-#| label: filtering 2022 data
+#| label: gss 2022
 gss2022 <- gss %>% 
@@
 
+view(gss2022)
 ```
+
+I noticed that there are 3544 entries, but the variable "id" goes to 3545.
+After investigation, I found that "id == 3419" doesn't exist and jumps from 3418 to 3420.
+It doesn't affect our work at the moment, so I'll leave it for now (and fix it later when needed).
+"id" is a functional unique identifier in our data.
+
+# analytical dataset
+
+Below, we create the analytical dataset.
+We use the gss2022 subset we created above, leaving the original dataset untouched -- this is important for replication purposes.
+First, we dropped the missing values for our dependent variable.
+The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
+For the purposes of this project, we are considering all of the above as missing values.
+
+Then we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This was not going to work.
+So, we opted to use the "racerank1" variable, since in this case participants were able to answer with either a race or ethnicity.
+There were 16 different responses: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
+We recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
+We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
+Additionally, because race is an important control that we identify, we dropped the missing values.
 
@@
 data <- gss2022 %>% 
-    select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
+    select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
+
+# instead of using view(), I would suggest using head() to take a look at the data within quarto.  
+head(data)
+view(data)
 
 ```
-
-The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
-For the purposes of this project, we are considering all of the above as missing values.
-Below we recode all these indicators as NAs.
-This is necessary to run mice imputations to address missing values.
 
@@
 na_data <- data %>% 
-  mutate(across(c(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
+  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
                 ~ifelse(. %in% missing_vec, NA_character_, .)))
+
+# instead of using view(), I would suggest using head() to take a look at the data within quarto.  
+head(na_data)
+view(na_data)
 
@@
 
-To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
-
 ```{r}
 #| label: mice imputation
-# DO NOT RUN 
+# I am unsure what this code chunk does. A comment may be useful for reproducibility.
+
 complete5 <- mice(na_data, 
-                 m = 50, 
+                 m=50, 
                  seed = 25)
 
-summary(complete(complete5, 1))
+complete(complete5,1)
 
-temp <- complete(complete5, action = "long", include = TRUE)
+temp <- complete(complete5, action="long", include=TRUE)
 
@@
 
-save(tidydata_complete, file = "data/tidydata_complete.RData")
-
 ```
 
-Fear is our dependent variable, so the first step was to remove NAs for that variabble.
-
-Race/ ethnicity is an important control for our models.
-First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
-So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
-We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
-We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
-
 ```{r}
-#| label: recoding variables 
+#| label: recode-variables
 tidydata <- na_data %>% 
@@
   mutate(racecombo = 
-           factor(case_when(racerank1 == "White" ~ "White",
-                     racerank1 == "Black or African-American" ~ "Black",
+           case_when(racerank1 == "White" ~ "White",
+                     racerank1 == "Black or African American" ~ "Black",
                      racerank1 == "Hispanic" ~ "Hispanic",
@@
                        racerank1 == "Other Pacific Islander" ~ "NHPI",
-                      racerank1 == "Some other race" ~ "Other"))) %>% 
-  mutate(racecombo = fct_relevel(racecombo, 
-              "White",
-              "Black", 
-              "Asian",
-              "Hispanic",
-              "AIAN",
-              "NHPI",
-              "Other")) %>%   
-  mutate(partnrs5 = 
-           factor(case_when(partnrs5 == "NO PARTNERS" ~ "0",
-                            partnrs5 == "1 PARTNER" | partnrs5 == "2 PARTNERS" |
-                              partnrs5 == "3 PARTNERS" | 
-                              partnrs5 == "4 PARTNERS" ~ "Less than 5",
-                            partnrs5 == "5-10 PARTNERS" ~ "5-10",
-                            partnrs5 == "11-20 PARTNERS" ~ "11-20",
-                            partnrs5 == "21-100 PARTNERS" ~ "21-100",
-                            partnrs5 == "MORE THAN 100 PARTNERS" ~ "100 or more",
-                            .default = NA))) %>%
-  mutate(partnrs5 = fct_relevel(partnrs5,
-                                "0",
-                                "Less than 5",
-                                "5-10",
-                                "11-20",
-                                "21-100",
-                                "100 or more")) %>%
-  mutate(uscitzn = 
-           factor(case_when(uscitzn == "A u.s. citizen" | uscitzn == "A u.s. citizen born in puerto rico, the u.s. virgin islands, or the northern marianas islands (if volunteered)" |
-                       uscitzn == "Born outside of the u.s. to parents who were u.s. citizens at that time (if volunteered)" ~ "Yes",
-                     uscitzn == "Not a u.s. citizen" ~ "No",
-                     .default = NA))) %>% 
-   mutate(dwelown =
-           factor(case_when(dwelown == "Own or is buying" ~ "Owner",
-                            dwelown == "Pays rent" ~ "Renter",
-                            dwelown == "Other" ~ "Other"))) %>% 
-  mutate(degree = fct_relevel(degree, 
-                              "Less than high school",
-                              "High school",
-                              "Associate/junior college",
-                              "Bachelor's",
-                              "Graduate")) %>% 
-  mutate(attend = fct_relevel(attend, 
-                              "Never",
-                              "Less than once a year",
-                              "About once or twice a year",
-                              "2-3 times a month",
-                              "Nearly every week",
-                              "Every week",
-                              "Several times a week")) %>% 
-  mutate(happy = fct_relevel(happy, 
-                             "Not too happy",
-                             "Pretty happy",
-                             "Very happy",)) %>% 
-  mutate(health = fct_relevel(health, 
-                              "Poor",
-                              "Fair",
-                              "Good",
-                              "Excellent")) %>% 
-  mutate(helpful = fct_relevel(helpful, 
-                               "Looking out for themselves",
-                               "Try to be helpful",
-                               "Depends")) %>% 
-  mutate(fair = fct_relevel(fair, 
-                            "Would take advantage of you",
-                            "Would try to be fair",
-                            "Depends")) %>% 
-  mutate(trust = fct_relevel(trust, 
-                             "Can't be too careful",
-                             "Can trust",
-                             "Depends")) %>% 
-  select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
+                      racerank1 == "Some other race" ~ "Other")) %>% 
+  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")
 
+#the above recoding can also be done using factors. I also noticed that the results were difficult.
+
+na_data %>%
+  mutate(racecombo1 = fct_recode(racerank1,
+                                "White" = "White",
+                                "Black or African American" = "Black",
+                                "Hispanic" = "Hispanic",
+                                "Asian" = "Asian Indian",
+                                "Asian" = "Filipino",
+                                "Asian" = "Japanese",
+                                "Asian" = "Korean",
+                                "Asian" = "Vietnamese",
+                                "Asian" = "Other Asian",
+                                "AIAN" = "American Indian or Alaska Native",
+                                "NHPI" ="Guamanian or Chamorro",
+                                "NHPI" = "Samoan",
+                                "NHPI" = "Other Pacific Islander",
+                                "Other" = "Some other race")) %>%
+  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo1", "uscitzn", "dwelown") %>%
+  skim()
+
+# It might be a good idea to do some descriptive statistics here just to have a quick look at your data. I'm using skimr here to inspect the data.
+
+library(skimr)
+skim(na_data)
 ```
-
-```{r}
-```

### 2023-11-29 scripts/manuscript.qmd
eval:false for mice imputation
more commits

--- 2023-11-28 scripts/manuscript.qmd
@@
 
-To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
-
-```{r}
-#| label: mice imputation
-#| eval: false
-complete5 <- mice(na_data, 
-                 m = 5, 
-                 seed = 25)
-
-summary(complete(complete5, 1))
-
-temp <- complete(complete5, action = "long", include = TRUE)
-
-tidydata_complete <- as.mids(temp)
-
-save(tidydata_complete, file = "data/tidydata_complete.RData")
-
-```
-
 Fear is our dependent variable, so the first step was to remove NAs for that variabble.
@@
 
+To deal with the large amount of missing values, we used mice imputations. In the code below we create a new dataset with the imputed values. We will use this dataset in the future to run models. 
+```{r}
+#| label: mice imputation
+#| eval: false
+
+complete5 <- mice(gss_subset, 
+                 m = 5, 
+                 seed = 25)
+
+summary(complete(complete5, 1))
+
+temp5 <- complete(complete5, action = "long", include = TRUE)
+
+tidydata_complete <- as.mids(temp5)
+
+save(tidydata_complete, file = "data/tidydata_complete.RData")
+
+```
+
 ```{r}
@@
 gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
   pivot_longer(

### 2023-11-29 scripts/manuscript.qmd
Added some explanatory texts

--- 2023-11-28 scripts/manuscript.qmd
@@
 
-To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
+Fear is our dependent variable, so the first step was to remove observations with NAs for that variable, as those won't be meaningful for our analysis.
 
-```{r}
-#| label: mice imputation
-#| eval: false
-complete5 <- mice(na_data, 
-                 m = 5, 
-                 seed = 25)
-
-summary(complete(complete5, 1))
-
-temp <- complete(complete5, action = "long", include = TRUE)
-
-tidydata_complete <- as.mids(temp)
-
-save(tidydata_complete, file = "data/tidydata_complete.RData")
-
-```
-
-Fear is our dependent variable, so the first step was to remove NAs for that variabble.
-
-Race/ ethnicity is an important control for our models.
+Informed by theory, race/ ethnicity is an important covariate for our models.
 First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
@@
 gss_subset <- tidydata %>% 
-  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
-         partnrs5, fear) 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
+```
+
+To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
+
+```{r}
+#| label: mice imputation
+
+complete5 <- mice(gss_subset, 
+                 m = 5, 
+                 seed = 25)
+
+summary(complete(complete5, 1))
+
+temp5 <- complete(complete5, action = "long", include = TRUE)
+
+tidydata_complete <- as.mids(temp5)
+
+save(tidydata_complete, file = "data/tidydata_complete.RData")
+
 ```
@@
 
+I want to make a note about the interpretation for "mean fear by race".
+Because our outcome variable is dichotomous (yes or no to having fear of walking alone in neighborhood at night).
+The mean number approximately gets at the idea of the proportions of a specific racial group that experience fear of walking in neighborhood at night.
+
 ```{r}

### 2023-11-29 scripts/manuscript.qmd
2 tables and 2 figures on gender&fear

@@
 library(ggplot2)
+library(vcd)
 ```
@@
 
+# Tables and Figures
+
+Let's understand our sample a little better. There are 2318 observations and 11 selected variables in our data set.
+
+```{r}
+#| label: descriptive statistics
+view(gss_subset)
+
+skimr::skim(gss_subset)
+
+```
+
+```{r}
+#| label: table - gender and fear
+gss_subset %>% 
+  tabyl(fear, sexnow1) %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns(position = "front") %>% 
+  adorn_title(
+    row_name = "Fear at Night",
+    col_name = "Gender",
+    placement = "combined"
+  ) %>% 
+  flextable::flextable() %>% 
+  flextable::autofit()
+```
+
+```{r}
+#| label: table - racecombo and fear
+gss_subset %>% 
+  tabyl(fear, racecombo) %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns(position = "front") %>% 
+  adorn_title(
+    row_name = "Fear at Night",
+    col_name = "Race",
+    placement = "combined"
+  ) %>% 
+  flextable::flextable() %>% 
+  flextable::autofit()
+```
+
+```{r}
+#| label: fig1-fear_gender
+#| fig-cap: Bar graph showing the proportion or respondants, by gender, that feared walking alone at night.
+
+gss_subset %>% 
+  ggplot(aes(x = sexnow1, y = after_stat(prop), group = fear, 
+             fill = fear)) +
+  geom_bar(position = "dodge") + 
+  scale_y_continuous(labels = scales::percent) + 
+  labs(title = "Social Fear by Gender",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Gender",
+       y = NULL,
+       fill = "Fear"
+       ) +
+  theme_bw()
+
+```
+
+```{r}
+#| label: fig2-fear_gender
+#| fig-cap: Bar graph showing the proportion or respondants, by gender, that feared walking alone at night.
+
+gss_subset %>% 
+  count(sexnow1, fear) %>% 
+  ggplot() +
+  geom_col(
+    mapping = aes(x = fear,
+                  fill = sexnow1,
+                  y = n)
+  ) +
+  coord_flip() +
+  labs(title = "Social Fear by Gender",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Count",
+       y = "Fear",
+       )
+  theme_bw()
+```
+
 ```{r}

### 2023-11-29 scripts/manuscript.qmd
made two summary tables

@@
 
-complete5 <- mice(gss_subset, 
-                 m = 5, 
-                 seed = 25)
+#complete5 <- mice(gss_subset, 
+ #                m = 5, 
+  #               seed = 25)
 
-summary(complete(complete5, 1))
+#summary(complete(complete5, 1))
 
-temp5 <- complete(complete5, action = "long", include = TRUE)
+#temp5 <- complete(complete5, action = "long", include = TRUE)
 
-tidydata_complete <- as.mids(temp5)
+#tidydata_complete <- as.mids(temp5)
 
-save(tidydata_complete, file = "data/tidydata_complete.RData")
+#save(tidydata_complete, file = "data/tidydata_complete.RData")
 
@@
 ```{r}
-#| label: descriptive statistics
+#| label: get to know data more
 view(gss_subset)
@@
 skimr::skim(gss_subset)
-
 ```
@@
 ```{r}
-#| label: table - gender and fear
+#| label: descriptive statistics
+gss_subset %>% 
+  select(fear, sexnow1, racecombo, degree) %>% 
+  gtsummary::tbl_summary()
+```
+
+```{r}
+#| label: tbl_gender_fear
+#| tbl-cap: table showing the distribution of fear by gender
+
 gss_subset %>% 
@@
 ```{r}
-#| label: table - racecombo and fear
+#| label: tbl_racecombo_fear
+#| tbl-cap: table showing the distribution of fear by race
 gss_subset %>% 
@@
 #| fig-cap: Bar graph showing the proportion or respondants, by gender, that feared walking alone at night.
+#| message: false
 

### 2023-11-29 scripts/manuscript.qmd
Comments

--- 2023-11-29 scripts/manuscript.qmd
@@
 
-# Introduction
+# codebook for variables
 
-# Methods
+**Dependent Variable** "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
-# Results
+**Sociodemographics Variables** 1.
+"born": Were you born in this country?
+2.
+"age" 3.
+"sexnow1": Do you describe yourself as male, female, or transgender?
+*This is self-described gender.* 4.
+"sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.* 5.
+"sexornt": Which of the following best describes you?
+6.
+"racerank1": If you had to choose, which of these races do you identify most with?
+*Kept this race variable here instead of "race" because it allows more heterogeneity within group.* 7.
+"hispanic": Are you Spanish, Hispanic, or Latino/Latina?
+IF YES: Which group are you from?
+8.
+"degree": Respondent's degree.
+9.
+"uscitzn": Are you a US citizen?
+10.
+"dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
-# Discussion
+**Other Controls** 1 "attend": How often do you attend religious services?
+2.
+"raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+3.
+"happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+4.
+"health": Would you say your own health, in general, is excellent, good, fair, or poor?
+5.
+"helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+6.
+"fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+7.
+"trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+8.
+"partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
 
-# Codebook
+***Thank you for providing the detail for your variables***
 
-**Dependent Variable**
-
--   "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
-
-**Sociodemographics Variables**
-
--   "born": Were you born in this country?
-
--   "age"
-
--   "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
-
--   "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
-
--   "sexornt": Which of the following best describes you?
-
--   "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
-
--   "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
-
--   "degree": Respondent's degree.
-
--   "uscitzn": Are you a US citizen?
-
--   "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
-
-**Other Controls**
-
--   "attend": How often do you attend religious services?
-
--   "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
-
--   "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
-
--   "health": Would you say your own health, in general, is excellent, good, fair, or poor?
-
--   "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
-
--   "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
-
--   "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
-
--   "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
-
-# Data Cleaning
+# set up
 
@@
 library(mice)
-library(forcats)
-library(ggplot2)
+
+# I don't think I've seen the mice package before. Curious to see how you use that.
 ```
@@
 #| label: loading data
+
 gss <- import(here("data", "GSS.xlsx")) %>%
   as_tibble() %>% 
+  characterize() %>% # might not need characterize, TBD
   clean_names()
 
+view(gss)
+
+# Great use of notes
 ```
 
+For the purposes of EDLD 651 final project, we only use data from the year 2022. This will result in a much smaller data set, and we won't need to work with panel models for now.
+
 ```{r}
-#| label: filtering 2022 data
+#| label: gss 2022
 gss2022 <- gss %>% 
@@
 data <- gss2022 %>% 
-    select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
+    select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)
 
+view(data)
 ```
-
-The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
-For the purposes of this project, we are considering all of the above as missing values.
-Below we recode all these indicators as NAs.
-This is necessary to run mice imputations to address missing values.
 
@@
 na_data <- data %>% 
-  mutate(across(c(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
+  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
                 ~ifelse(. %in% missing_vec, NA_character_, .)))
+
+view(na_data)
+
+# Cool code
+```
+
+```{r}
+#| label: mice imputation
+
+complete5 <- mice(na_data, 
+                 m=50, 
+                 seed = 25)
+
+complete(complete5,1)
+
+temp <- complete(complete5, action="long", include=TRUE)
+
+tidydata_complete <- as.mids(temp)
 
@@
 
-Fear is our dependent variable, so the first step was to remove observations with NAs for that variable, as those won't be meaningful for our analysis.
-
-Informed by theory, race/ ethnicity is an important covariate for our models.
-First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
-So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
-We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
-We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
-
 ```{r}
-#| label: recoding variables 
+#| label: recode-variables
 tidydata <- na_data %>% 
   filter(!is.na(fear)) %>% 
-  mutate(fear = factor(case_when(fear == "NO" ~ "no",
-                                 fear == "YES" ~ "yes"))) %>% 
   mutate(racecombo = 
-           factor(case_when(racerank1 == "White" ~ "White",
-                     racerank1 == "Black or African-American" ~ "Black",
+           case_when(racerank1 == "White" ~ "White",
+                     racerank1 == "Black or African American" ~ "Black",
                      racerank1 == "Hispanic" ~ "Hispanic",
@@
                        racerank1 == "Other Pacific Islander" ~ "NHPI",
-                      racerank1 == "Some other race" ~ "Other"))) %>% 
-  mutate(racecombo = fct_relevel(racecombo, 
-              "White",
-              "Black", 
-              "Asian",
-              "Hispanic",
-              "AIAN",
-              "NHPI",
-              "Other")) %>%   
-  mutate(partnrs5 = 
-           factor(case_when(partnrs5 == "NO PARTNERS" ~ "0",
-                            partnrs5 == "1 PARTNER" | partnrs5 == "2 PARTNERS" |
-                              partnrs5 == "3 PARTNERS" | 
-                              partnrs5 == "4 PARTNERS" ~ "Less than 5",
-                            partnrs5 == "5-10 PARTNERS" ~ "5-10",
-                            partnrs5 == "11-20 PARTNERS" ~ "11-20",
-                            partnrs5 == "21-100 PARTNERS" ~ "21-100",
-                            partnrs5 == "MORE THAN 100 PARTNERS" ~ "100 or more",
-                            .default = NA))) %>%
-  mutate(partnrs5 = fct_relevel(partnrs5,
-                                "0",
-                                "Less than 5",
-                                "5-10",
-                                "11-20",
-                                "21-100",
-                                "100 or more")) %>%
-  mutate(uscitzn = 
-           factor(case_when(uscitzn == "A u.s. citizen" | uscitzn == "A u.s. citizen born in puerto rico, the u.s. virgin islands, or the northern marianas islands (if volunteered)" |
-                       uscitzn == "Born outside of the u.s. to parents who were u.s. citizens at that time (if volunteered)" ~ "Yes",
-                     uscitzn == "Not a u.s. citizen" ~ "No",
-                     .default = NA))) %>% 
-   mutate(dwelown =
-           factor(case_when(dwelown == "Own or is buying" ~ "Owner",
-                            dwelown == "Pays rent" ~ "Renter",
-                            dwelown == "Other" ~ "Other"))) %>% 
-  mutate(degree = fct_relevel(degree, 
-                              "Less than high school",
-                              "High school",
-                              "Associate/junior college",
-                              "Bachelor's",
-                              "Graduate")) %>% 
-  mutate(attend = fct_relevel(attend, 
-                              "Never",
-                              "Less than once a year",
-                              "About once or twice a year",
-                              "2-3 times a month",
-                              "Nearly every week",
-                              "Every week",
-                              "Several times a week")) %>% 
-  mutate(happy = fct_relevel(happy, 
-                             "Not too happy",
-                             "Pretty happy",
-                             "Very happy",)) %>% 
-  mutate(health = fct_relevel(health, 
-                              "Poor",
-                              "Fair",
-                              "Good",
-                              "Excellent")) %>% 
-  mutate(helpful = fct_relevel(helpful, 
-                               "Looking out for themselves",
-                               "Try to be helpful",
-                               "Depends")) %>% 
-  mutate(fair = fct_relevel(fair, 
-                            "Would take advantage of you",
-                            "Would try to be fair",
-                            "Depends")) %>% 
-  mutate(trust = fct_relevel(trust, 
-                             "Can't be too careful",
-                             "Can trust",
-                             "Depends")) %>% 
-  select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
+                      racerank1 == "Some other race" ~ "Other")) %>% 
+  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")
 
 ```
-
-```{r}
-#| label: creating a smaller subset from data
-
-gss_subset <- tidydata %>% 
-  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
-```
-
-To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
-
-```{r}
-#| label: mice imputation
-
-complete5 <- mice(gss_subset, 
-                 m = 5, 
-                 seed = 25)
-
-summary(complete(complete5, 1))
-
-temp5 <- complete(complete5, action = "long", include = TRUE)
-
-tidydata_complete <- as.mids(temp5)
-
-save(tidydata_complete, file = "data/tidydata_complete.RData")
-
-```
-
-```{r}
-#| label: tbl-fear_race
-#| tbl-cap: Table showing mean fear by race.
-
-gss_subset %>% 
-  pivot_longer(
-    cols = fear,
-    names_to = "fear",
-    values_to = "answer") %>% 
-  select(racecombo, answer) %>% 
-  mutate(fear = as.numeric(case_when(answer == "no" ~ "0",
-                                 answer == "yes" ~ "1"))) %>% 
-  select(racecombo, fear) %>% 
-  group_by(racecombo) %>% 
-  summarize(mean_fear = mean(fear)) %>% 
-  kable(caption = "Mean Social Fear by Race",
-        col.names = c("Race", "Mean"),
-        digits = 3)
-```
-
-I want to make a note about the interpretation for "mean fear by race".
-Because our outcome variable is dichotomous (yes or no to having fear of walking alone in neighborhood at night).
-The mean number approximately gets at the idea of the proportions of a specific racial group that experience fear of walking in neighborhood at night.
-
-```{r}
-#| label: fig-fear_race
-#| fig-cap: Bar graph showing the proportion or respondants, by race, that feared walking alone at night.
-
-gss_subset %>% 
-  filter(!is.na(racecombo)) %>% 
-  filter(racecombo != "Other") %>% 
-  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
-             fill = fear)) +
-  geom_bar(position = "dodge") + 
-  scale_y_continuous(labels = scales::percent) +
-  labs(title = "Social Fear by Race",
-       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
-       x = "Race",
-       y = NULL,
-       fill = "Fear"
-       ) +
-  theme_bw()
-
-```

### 2023-11-29 scripts/manuscript.qmd
pretty pictures

--- 2023-11-29 scripts/manuscript.qmd
@@
 library(ggplot2)
-library(vcd)
+library(RColorBrewer)
 ```
@@
 
-Fear is our dependent variable, so the first step was to remove observations with NAs for that variable, as those won't be meaningful for our analysis.
+Fear is our dependent variable, so the first step was to remove NAs for that variabble.
 
-Informed by theory, race/ ethnicity is an important covariate for our models.
+Race/ ethnicity is an important control for our models.
 First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
@@
   filter(!is.na(fear)) %>% 
+  mutate(age = as.numeric(age)) %>% 
+  mutate(raclive = as.factor(raclive)) %>% 
+  mutate(sexnow1 = as.factor(sexnow1)) %>% 
   mutate(fear = factor(case_when(fear == "NO" ~ "no",
@@
 gss_subset <- tidydata %>% 
-  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
+         partnrs5, fear) 
 ```
 
-To deal with the large amount of missing values, we used mice imputations. Currently, there are errors in my code, so this is not running correctly. Feedback welcomed.
+To deal with the large amount of missing values, we used mice imputations. In the code below we create a new dataset with the imputed values. We will use this dataset in the future to run models.
 
@@
 #| label: mice imputation
+#| eval: false
 
@@
 gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  filter(!is.na(sexnow1)) %>% 
+  filter(sexnow1 != "None of these") %>% 
   pivot_longer(
@@
     values_to = "answer") %>% 
-  select(racecombo, answer) %>% 
+  select(racecombo, answer, sexnow1) %>% 
   mutate(fear = as.numeric(case_when(answer == "no" ~ "0",
                                  answer == "yes" ~ "1"))) %>% 
-  select(racecombo, fear) %>% 
-  group_by(racecombo) %>% 
+  select(racecombo, fear, sexnow1) %>% 
+  group_by(racecombo, sexnow1) %>% 
   summarize(mean_fear = mean(fear)) %>% 
-  kable(caption = "Mean Social Fear by Race",
-        col.names = c("Race", "Mean"),
+  pivot_wider(
+    names_from = sexnow1,
+    values_from = mean_fear
+  ) %>% 
+  kable(caption = "Social Fear by Race and Gender",
+        col.names = c("Race", "Female", "Male", "Transgender"),
         digits = 3)
-```
 
-I want to make a note about the interpretation for "mean fear by race".
-Because our outcome variable is dichotomous (yes or no to having fear of walking alone in neighborhood at night).
-The mean number approximately gets at the idea of the proportions of a specific racial group that experience fear of walking in neighborhood at night.
-
-# Tables and Figures
-
-Let's understand our sample a little better. There are 2318 observations and 11 selected variables in our data set.
-
-```{r}
-#| label: descriptive statistics
-view(gss_subset)
-
-skimr::skim(gss_subset)
-
+# to complete later; kableextra to add n and other details 
 ```
@@
 ```{r}
-#| label: table - gender and fear
-gss_subset %>% 
-  tabyl(fear, sexnow1) %>% 
-  adorn_percentages() %>% 
-  adorn_pct_formatting() %>% 
-  adorn_ns(position = "front") %>% 
-  adorn_title(
-    row_name = "Fear at Night",
-    col_name = "Gender",
-    placement = "combined"
-  ) %>% 
-  flextable::flextable() %>% 
-  flextable::autofit()
-```
-
-```{r}
-#| label: table - racecombo and fear
-gss_subset %>% 
-  tabyl(fear, racecombo) %>% 
-  adorn_percentages() %>% 
-  adorn_pct_formatting() %>% 
-  adorn_ns(position = "front") %>% 
-  adorn_title(
-    row_name = "Fear at Night",
-    col_name = "Race",
-    placement = "combined"
-  ) %>% 
-  flextable::flextable() %>% 
-  flextable::autofit()
-```
-
-```{r}
-#| label: fig1-fear_gender
-#| fig-cap: Bar graph showing the proportion or respondants, by gender, that feared walking alone at night.
-
-gss_subset %>% 
-  ggplot(aes(x = sexnow1, y = after_stat(prop), group = fear, 
-             fill = fear)) +
-  geom_bar(position = "dodge") + 
-  scale_y_continuous(labels = scales::percent) + 
-  labs(title = "Social Fear by Gender",
-       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
-       x = "Gender",
-       y = NULL,
-       fill = "Fear"
-       ) +
-  theme_bw()
-
-```
-
-```{r}
-#| label: fig2-fear_gender
-#| fig-cap: Bar graph showing the proportion or respondants, by gender, that feared walking alone at night.
-
-gss_subset %>% 
-  count(sexnow1, fear) %>% 
-  ggplot() +
-  geom_col(
-    mapping = aes(x = fear,
-                  fill = sexnow1,
-                  y = n)
-  ) +
-  coord_flip() +
-  labs(title = "Social Fear by Gender",
-       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
-       x = "Count",
-       y = "Fear",
-       )
-  theme_bw()
-```
-
-```{r}
-#| label: fig-fear_race
+#| label: fig-FearRace
 #| fig-cap: Bar graph showing the proportion or respondants, by race, that feared walking alone at night.
@@
   filter(racecombo != "Other") %>% 
-  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
-             fill = fear)) +
-  geom_bar(position = "dodge") + 
+  group_by(racecombo) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = racecombo, y = prop, fill = racecombo)) +
+  geom_col() + 
+  scale_fill_brewer(palette = "Accent") +
   scale_y_continuous(labels = scales::percent) +
   labs(title = "Social Fear by Race",
-       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       caption = "Question: Is there any area right around here where you would be afraid to walk alone at night?",
        x = "Race",
-       y = NULL,
-       fill = "Fear"
-       ) +
-  theme_bw()
+       y = NULL) +
+  theme_bw() +
+  guides(fill = "none")
 

### 2023-11-29 scripts/manuscript.qmd
github merge test
test 2
tables and figures progress from Nov 28 evening

@@
 
-Race/ ethnicity is an important control for our models.
+Informed by theory, race/ ethnicity is an important covariate for our models.
 First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
 So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
+We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
+We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
+
+Fear is our dependent variable, so the first step was to remove NAs for that variabble.
+
+Race/ ethnicity is an important control for our models.
+First, we sought to create a combined race variable from the variables “race” and “hispanic.” After looking at the data, we saw that the options for the race variable were “Black, White, Other.” This does not provide the control we need.
+So, we opted to use the “racerank1” variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
 We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
@@
 gss_subset <- tidydata %>% 
-  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
-         partnrs5, fear) 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
 ```
@@
 ```
+
+# Tables and Figures
+
+```{r}
+view(gss_subset)
+
+skimr::skim(gss_subset)
+```
+
+```{r}
+gss_subset %>% 
+  select(fear, sexnow1, racecombo1, degree) %>% 
+  gtsummary::tbl_summary()
+```
+
+```{r}
+gss_subset %>% 
+  tabyl(fear, sexnow1) %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns(position = "front") %>% 
+  adorn_title(
+    row_name = "Fear at Night",
+    col_name = "Gender",
+    placement = "combined"
+  ) %>% 
+  flextable::flextable() %>% 
+  flextable::autofit()
+```
+
+```{r}
+gss_subset %>% 
+  tabyl(fear, racecombo) %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns(position = "front") %>% 
+  adorn_title(
+    row_name = "Fear at Night",
+    col_name = "Race",
+    placement = "combined"
+  ) %>% 
+  flextable::flextable() %>% 
+  flextable::autofit()
+```
+
+```{r}
+gss_subset %>% 
+  ggplot(aes(x = sexnow1, y = after_stat(prop), group = fear, 
+             fill = fear)) +
+  geom_bar(position = "dodge") + 
+  scale_y_continuous(labels = scales::percent) + 
+  labs(title = "Social Fear by Gender",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Gender",
+       y = NULL,
+       fill = "Fear"
+       ) +
+  theme_bw()
+```
+
+```{r}
+gss_subset %>% 
+  count(sexnow1, fear) %>% 
+  ggplot() +
+  geom_col(
+    mapping = aes(x = fear,
+                  fill = sexnow1,
+                  y = n)
+  ) +
+  coord_flip() +
+  labs(title = "Social Fear by Gender",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Count",
+       y = "Fear",
+       )+
+  theme_bw()
+```
+
+```{r}
+gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
+             fill = fear)) +
+  geom_bar(position = "dodge") + 
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Race",
+       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
+       x = "Race",
+       y = NULL,
+       fill = "Fear"
+       ) +
+  theme_bw()
+```

### 2023-12-02 scripts/manuscript.qmd
Figures complete

--- 2023-11-29 scripts/manuscript.qmd
@@
 format: html
+toc: true
+toc-location: left
+toc-title: Table of Contents
 highlight-style: pygments
@@
   warning: false
+fig-width: 6.5
+fig-height: 8
 code-fold: true
-code-summary: "Show the code"
+code-summary: "View Code"
 embed-resources: true
 bibliography: references.bib
+link-citations: true
 ---
@@
 
-# Codebook
+\newpage
 
-**Dependent Variable**
+# References
 
--   "fear": Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+\newpage
 
-**Sociodemographics Variables**
+# Appendices
 
--   "born": Were you born in this country?
+## Appendix 1: Codebook
 
--   "age"
+### Dependent Variable
 
--   "sexnow1": Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+-   **fear**: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
--   "sex": Code respondent's sex *This is gender presentation as interpreted by the interviewer.*
+### Independent Variables
 
--   "sexornt": Which of the following best describes you?
+#### Individual-Level
 
--   "racerank1": If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
+-   **born**: Were you born in this country?
 
--   "hispanic": Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
+-   **age**
 
--   "degree": Respondent's degree.
+-   **sexnow1**: Do you describe yourself as male, female, or transgender? *This is self-described gender.*
 
--   "uscitzn": Are you a US citizen?
+-   **sex**: Code respondent's sex (*This is gender presentation as interpreted by the interviewer).*
 
--   "dwelown": (Do you/Does your family) own your (home/apartment), pay rent, or what?
+-   **sexornt**: Which of the following best describes you?
 
-**Other Controls**
+-   **racerank1**: If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
 
--   "attend": How often do you attend religious services?
+-   **hispanic**: Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?
 
--   "raclive": Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+-   **degree**: Respondent's degree.
 
--   "happy": Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
+-   **uscitzn**: Are you a US citizen?
 
--   "health": Would you say your own health, in general, is excellent, good, fair, or poor?
+-   **dwelown**: (Do you/Does your family) own your (home/apartment), pay rent, or what?
 
--   "helpful": Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+#### Community-Level
 
--   "fair": Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+-   **attend**: How often do you attend religious services?
 
--   "trust": Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+-   **raclive**: Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
 
--   "partnrs5": since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+-   **happy**: Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?
 
-# Data Cleaning
+-   **health**: Would you say your own health, in general, is excellent, good, fair, or poor?
+
+-   **helpful**: Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?
+
+-   **fair**: Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?
+
+-   **trust**: Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?
+
+-   **partnrs5**: since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
+
+## Appendix II: Data Cleaning
 
@@
 ```{r}
-#| label: loading data 
+#| label: load-data
+
 gss <- import(here("data", "GSS.xlsx")) %>%
@@
 ```{r}
-#| label: filtering 2022 data
+#| label: filtering-2022-data
+
 gss2022 <- gss %>% 
@@
 #| label: select-variables
+
 data <- gss2022 %>% 
@@
 For the purposes of this project, we are considering all of the above as missing values.
-Below we recode all these indicators as NAs.
-This is necessary to run mice imputations to address missing values.
+For our next steps, we need to reexamine this decision as".i: Inapplicable" is a valid skip and needs to be accounted for accordingly.
+
+In the following code chunk, we recode the aforementioned indicators as NAs. This is a necessary first step to run mice imputations.
 
 ```{r}
-#| label: coding all missing data as NA
+#| label: coding-missing-data-as-NAs
 
@@
 
-na_data <- data %>% 
+tidydata <- data %>% 
   mutate(across(c(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
@@
 
-Fear is our dependent variable, so the first step was to remove NAs for that variabble.
+First we filtered out all NAs from our dependent variable, fear.
 
-Race/ ethnicity is an important control for our models.
+Informed by theory, race/ ethnicity is an important covariate for our models.
 First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
@@
 
+Then we factored and recoded selected variables for our analysis. It was important to not have any variables as character strings, otherwise mice imputations would not work.
+
 ```{r}
-#| label: recoding variables 
-tidydata <- na_data %>% 
+#| label: factoring-recoding-variables
+
+tidydata <- tidydata %>% 
   filter(!is.na(fear)) %>% 
@@
                              "Can trust",
-                             "Depends")) %>% 
-  select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racecombo, uscitzn, dwelown)
+                             "Depends"))
 
@@
 
+Once our data was tidy, we selected the variables we would be using for our analysis.
+
 ```{r}
-#| label: creating a smaller subset from data
+#| label: subset-data
 
 gss_subset <- tidydata %>% 
-  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, 
-         partnrs5, fear) 
+  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
 ```
 
-To deal with the large amount of missing values, we used mice imputations. In the code below we create a new dataset with the imputed values. We will use this dataset in the future to run models.
+To deal with the large amount of missing values, we ran mice imputations.
+In the code below we create a new dataset with the imputed values.
+We will use this dataset in our next stage to run various models.
 
 ```{r}
-#| label: mice imputation
+#| label: mice-imputation
 #| eval: false
@@
 
+\newpage
+
+# Tables and Figures
+
 ```{r}
-#| label: tbl-fear_race
-#| tbl-cap: Table showing mean fear by race.
+#| label: tbl-FearRaceSex
+#| tbl-cap: Mean fear by race and gender calculated from respondents affirmative answer to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
@@
         col.names = c("Race", "Female", "Male", "Transgender"),
-        digits = 3)
-
-# to complete later; kableextra to add n and other details 
+        digits = 3) %>% 
+  kableExtra::row_spec(0:6, background = "white") 
 ```
@@
 ```{r}
-#| label: fig-FearRace
-#| fig-cap: Bar graph showing the proportion or respondants, by race, that feared walking alone at night.
+#| label: fig-FearRaceGender
+#| fig-cap: Bar graph showing the percentage of respondants, by race and gender, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
 
@@
   filter(racecombo != "Other") %>% 
-  group_by(racecombo) %>% 
+  filter(!is.na(sexnow1)) %>% 
+  filter(sexnow1 != "None of these") %>% 
+  group_by(racecombo, sexnow1) %>% 
   summarize(fear = sum(fear == "yes"), 
@@
             prop = fear/n) %>% 
-  ggplot(aes(x = racecombo, y = prop, fill = racecombo)) +
-  geom_col() + 
-  scale_fill_brewer(palette = "Accent") +
+  ggplot(aes(x = racecombo, y = prop)) +
+  geom_col(aes(fill = sexnow1), position = "dodge") + 
+  scale_fill_brewer(palette = "Purples") +
   scale_y_continuous(labels = scales::percent) +
-  labs(title = "Social Fear by Race",
-       caption = "Question: Is there any area right around here where you would be afraid to walk alone at night?",
+  labs(title = "Social Fear by Race and Gender",
+       subtitle = "Is social fear influenced by respondent's race and gender?",
        x = "Race",
-       y = NULL) +
+       y = "Percentage of Respondents who Expressed Fear") +
   theme_bw() +
-  guides(fill = "none")
+  guides(fill = guide_legend(title = "Gender"))
 
 ```
+
+```{r}
+#| label: fig-FearRelgiousPart
+#| fig-cap: Bar graph showing the percentage of respondants, by degree of participation in religious activities, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(attend)) %>% 
+  group_by(attend) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = attend, y = prop, fill = attend)) +
+  geom_col() + 
+  scale_fill_brewer(palette = "Blues") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Relgious Participation",
+       subtitle = "Is social fear influenced by religious participation?",
+       x = "Religious Participation",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = "none") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
+
+```
+
+```{r}
+#| label: fig-FearHappiness
+#| fig-cap: Bar graph showing the percentage of respondants, by happiness levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(happy)) %>% 
+  group_by(happy) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = happy, y = prop, fill = happy)) +
+  geom_col() + 
+  scale_fill_brewer(palette = "PuBuGn") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Happiness Level",
+       subtitle = "Is social fear influenced by level of happiness?", 
+       x = "Happiness Level",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = "none") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
+```
+
+```{r}
+#| label: fig-FearNeighborhood
+#| fig-cap: Bar graph showing the percentage of respondants, based on race and neighborhood diversity, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  filter(!is.na(raclive)) %>%  
+  group_by(racecombo, raclive) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = racecombo, y = prop)) +
+  geom_col(aes(fill = raclive), position = "dodge") + 
+  scale_fill_brewer(palette = "Set2") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Race and Neighborhood Diversity",
+       subtitle = "Is social fear influenced by respondent's race and their neighborhood's diversity?",
+       x = "Race",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = guide_legend(title = "Residential Diversity")) 
+```
+
+```{r}
+#| label: fig-FearHealth
+#| fig-cap: Bar graph showing the percentage of respondants, by health levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(health)) %>% 
+  group_by(health) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = health, y = prop, fill = health)) +
+  geom_col() + 
+  scale_fill_brewer(palette = "PuRd") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Health Level",
+       subtitle = "Is social fear influenced by health?", 
+       x = "Health Level",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = "none") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
+```
+
+```{r}
+#| label: fig-FearPartnersGender
+#| fig-cap: Bar graph showing the percentage of respondants, by gender and number of sexual partners in the past 5 years, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(partnrs5)) %>% 
+  filter(!is.na(sexnow1)) %>% 
+  filter(sexnow1 != "None of these") %>% 
+  group_by(partnrs5, sexnow1) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = partnrs5, y = prop)) +
+  geom_col(aes(fill = sexnow1), position = "dodge") + 
+  scale_fill_brewer(palette = "Spectral") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Gender and Sexual Partners",
+       subtitle = "Is social fear influenced by respondent's gender and number of sexual partners?",
+       x = "Number of Sexual Partners in the Past 5 Years",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = guide_legend(title = "Gender"))
+
+```

### 2023-12-04 scripts/manuscript.qmd
figure out how to do word count in Quarto
methods

@@
 
+I'm leaving myself a note here to figure out how to do word count in a Quarto document.
+This will allow me to directly write in the qmd file and save the trouble of writing elsewhere like a word document and pasting things here.
+
 # Methods
+
+This study draws data from the General Social Survey (GSS).
+The GSS is a nationally representative interview-based survey that has collected longitudinal data since 1972.
+The GSS contains questions on demographics, behaviors, attitudes, civil liberaties, crime and violence, intergroup tolerance, morality, national spending priorities, psychological well-being, social mobility, stress, and traumatic events.
+Our study is limited to data collected during the 2022 wave and we draw primarily on demographic and well-being variables.
 

### 2023-12-05 scripts/manuscript.qmd
Methods write-up

@@
 ---
-title: "Social Conditions of Fear"
+title: "The Social Conditioning of Fear"
 author:
@@
 The GSS is a nationally representative interview-based survey that has collected longitudinal data since 1972.
-The GSS contains questions on demographics, behaviors, attitudes, civil liberaties, crime and violence, intergroup tolerance, morality, national spending priorities, psychological well-being, social mobility, stress, and traumatic events.
+The GSS contains questions on demographics, behaviors, attitudes, civil liberties, crime and violence, intergroup tolerance, morality, national spending priorities, psychological well-being, social mobility, stress, and traumatic events.
 Our study is limited to data collected during the 2022 wave and we draw primarily on demographic and well-being variables.
 
-# Results
+Our dependent variable is coded as **fear** by the GSS.
+Respondents were asked, "*Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night*?" During the 2022 wave, there were 3,544 answers coded as follows: 878 = yes, 1440 = no, and 1226 = NA.
+Notably, the GSS codes missing values with multiple indicators--".i: Inapplicable," ".n: No answer," ".d: Do not Know/Cannot Choose," and ".s: Skipped on Web." For the purposes of our preliminary study, we coded all missing data as NA.
+Furthermore, we applied fixed removal of NAs for our dependent variable and we ran mice imputations for all other variables to create a more robust dataset that is necessary for further modeling.
 
-# Discussion
-
-\newpage
-
-# References
-
-\newpage
-
-# Appendices
-
-## Appendix 1: Codebook
-
-### Dependent Variable
-
--   **fear**: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
-
-### Independent Variables
-
-#### Individual-Level
+Our control and independent variables are representative of two different dimensions we sought to capture. The first dimension is individual-level or identity characteristics:
 
@@
 
--   **sexnow1**: Do you describe yourself as male, female, or transgender? *This is self-described gender.*
+-   **sexnow1**: Do you describe yourself as male, female, or transgender?
 
--   **sex**: Code respondent's sex (*This is gender presentation as interpreted by the interviewer).*
+-   **sex**: Code respondent's sex
 
@@
 
--   **racerank1**: If you had to choose, which of these races do you identify most with? *Kept this race variable here instead of "race" because it allows more heterogeneity within group.*
+-   **racerank1**: If you had to choose, which of these races do you identify most with?
 
@@
 
-#### Community-Level
+Informed by theory, race is an important co-variate for our models as racialization serves to structure the daily lives of individuals based on where they fall on the Black-white spectrum, their opportunities, and their sense of belonging [@bonilla-silva2022; @omi2015].
+First, we sought to create a combined race variable from the variables **race** and **hispanic**.
+After looking at the data, we saw that **race** was coded as "Black, White, or Other." This does not provide the control we need as it omits those in the "racial middle" that are more likely to identify with a panethnic label and reject the Black-white binary \[ @mora2014; @obrien2008; [@rodriguez2000]\].
+Instead, we opted to use the **racerank1** variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other, Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
+We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
+We understand that these categories are problematic and contentious [@haney-lópez2006; @dowling2014], however they are the standard with quantitative data.
+
+Similarly, gender was a contentious variable to account for.
+Although the GSS has several questions aimed at capturing respondent's gender, we found them to be lacking in breathe.
+For example, the GSS codes **sex** within a Male/Female binary based solely on interviewer discretion.
+Ultimately, we opted to use **sexnow1** since it was the variable that most closely captured self-described gender.
+However, it remains problematic as respondents were only provided four choice: Male, Female, Transgender, and None of these.
+Out of 2306 cases, only 10 were coded as Transgender and 9 as None of these.
+
+Another important variable we sought to control for was citizenship status, since citizenship status and the associated stigma of illegality and threat of deportation have been found to increase fear in daily life [@golash-boza2015; @kline2019].
+However, GSS data is not conducive to controlling for this variable due to the large number of NAs.
+Out of 2318 cases, 2164 were coded as NAs.
+
+The second dimension we sought to capture was community-level characteristics, i.e., the social location of individuals. For this dimension, we drew on the following variables:
 
@@
 
-## Appendix II: Data Cleaning
+# Results
+
+# Discussion
+
+\newpage
+
+# Data Script
 
@@
 library(RColorBrewer)
+library(spelling)
 ```
@@
 
-The GSS codes missing data as either ".i: Inapplicable", ".n: No answer", ".d: Do not Know/Cannot Choose", and ".s: Skipped on Web.
-For the purposes of this project, we are considering all of the above as missing values.
-For our next steps, we need to reexamine this decision as".i: Inapplicable" is a valid skip and needs to be accounted for accordingly.
-
-In the following code chunk, we recode the aforementioned indicators as NAs. This is a necessary first step to run mice imputations.
-
 ```{r}
@@
 ```
-
-First we filtered out all NAs from our dependent variable, fear.
-
-Informed by theory, race/ ethnicity is an important covariate for our models.
-First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need.
-So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
-We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other.
-We understand that these categories are problematic and contentious, however they are the standard with quantitative data.
-
-Then we factored and recoded selected variables for our analysis. It was important to not have any variables as character strings, otherwise mice imputations would not work.
 
@@
 
-Once our data was tidy, we selected the variables we would be using for our analysis.
-
 ```{r}
@@
 ```
-
-To deal with the large amount of missing values, we ran mice imputations.
-In the code below we create a new dataset with the imputed values.
-We will use this dataset in our next stage to run various models.
 
@@
 ```
+
+\newpage
+
+# References

### 2023-12-05 scripts/manuscript.qmd
results section, rough draft

@@
 First, we sought to create a combined race variable from the variables **race** and **hispanic**.
-After looking at the data, we saw that **race** was coded as "Black, White, or Other." This does not provide the control we need as it omits those in the "racial middle" that are more likely to identify with a panethnic label and reject the Black-white binary \[ @mora2014; @obrien2008; [@rodriguez2000]\].
+After looking at the data, we saw that **race** was coded as "Black, White, or Other." This does not provide the control we need as it omits those in the "racial middle" that are more likely to identify with a panethnic label and reject the Black-white binary [@mora2014; @obrien2008; @rodriguez2000].
 Instead, we opted to use the **racerank1** variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other, Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic.
@@
 
-# Results
+These variables seek to account for respondent's social embeddedness and networks of support.
+Specifically, **raclive** is an important variable to consider as it accounts for neighborhood composition, in the absence of a neighborhood dissimilarity index.
+While this variable only looks at Black-white racial composition, it provides a good benchmark to examine broader racial antagonisms.
+Research has documented the associated fears white experience when Blacks or people of other racial backgrounds move into their neighborhood [@lacayo2017; @massey2020].
 
-# Discussion
-
-\newpage
-
-# Data Script
+## Data Script
 
@@
 ```{r}
+#| label: calculate-MeanFear
+
+overall_fear <- gss_subset %>% 
+  mutate(fear = case_when(fear == "no" ~ "0",
+                                 fear == "yes" ~ "1")) %>% 
+  select(fear)
+mean_fear <- mean(as.numeric(overall_fear$fear, na.rm=TRUE))
+
+```
+
+```{r}
 #| label: mice-imputation
@@
 
-\newpage
+# Results
 
-# Tables and Figures
+In the absence of running models, in this section we will discuss descriptive statistics through a series of tables and plots.
+We seek to show how the experience of fear differed across different measures of individual-level and community-level characteristics.
+To begin with, across our sample of 2318 cases, mean fear was calculated at `mean_fear`.
+
+First, we examined how fear is experienced across race and gender.
+As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents that identified as Male.
+Particularly, respondents that were Transgender expressed greater fear, but this may be indicative of sample representation.
+For example, only one respondent identified as both Hispanic and Transgender, and since they responded affirmative to the question, this finding is heavily skewed.
+When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.
 
@@
 ```
+@fig-FearRaceGender does a good job at plotting this relationship.
 
@@
 
+Moving on to measures of social embeddedness, we examined fear across degrees of religious participation, degree of happiness and health, residential diversity, and sexual activity.
+Regarding religious participation, we did not discern any clear pattern, as @fig-FearReligiousPart displays below.
+
 ```{r}
-#| label: fig-FearRelgiousPart
+#| label: fig-FearReligiousPart
 #| fig-cap: Bar graph showing the percentage of respondants, by degree of participation in religious activities, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
@@
 ```
+
+Next, looking at measures of Happiness and Health, we found that generally as happiness and health increase, social fear decreases. This is shown in @fig-FearHappiness and @fig-FearHealth.
 
@@
 ```{r}
+#| label: fig-FearHealth
+#| fig-cap: Bar graph showing the percentage of respondants, by health levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(health)) %>% 
+  group_by(health) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = health, y = prop, fill = health)) +
+  geom_col() + 
+  scale_fill_brewer(palette = "PuRd") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Health Level",
+       subtitle = "Is social fear influenced by health?", 
+       x = "Health Level",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = "none") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
+```
+
+Next, looking at residential composition and race combined displayed in @fig-FearNeighborhood, we do notice a pattern where respondents living in residentially more diverse areas were more likely to respond affirmative to the question of social fear.
+
+```{r}
 #| label: fig-FearNeighborhood
@@
 
-```{r}
-#| label: fig-FearHealth
-#| fig-cap: Bar graph showing the percentage of respondants, by health levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
-
-gss_subset %>% 
-  filter(!is.na(health)) %>% 
-  group_by(health) %>% 
-  summarize(fear = sum(fear == "yes"), 
-            n = n(),
-            prop = fear/n) %>% 
-  ggplot(aes(x = health, y = prop, fill = health)) +
-  geom_col() + 
-  scale_fill_brewer(palette = "PuRd") +
-  scale_y_continuous(labels = scales::percent) +
-  labs(title = "Social Fear by Health Level",
-       subtitle = "Is social fear influenced by health?", 
-       x = "Health Level",
-       y = "Percentage of Respondents who Expressed Fear") +
-  theme_bw() +
-  guides(fill = "none") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
-```
+Lastly, we looked at sexual activity as a marker of social fear as represented in @fig-FearPartnersGender.
 
@@
 ```
+# Discussion
 
 \newpage
-
 # References

### 2023-12-06 scripts/manuscript.qmd
edits to methods section

@@
 Our dependent variable is coded as **fear** by the GSS.
-Respondents were asked, "*Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night*?" During the 2022 wave, there were 3,544 answers coded as follows: 878 = yes, 1440 = no, and 1226 = NA.
+Respondents were asked, "*Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?*" During the 2022 wave, there were 3,544 answers coded as follows: 878 = yes, 1440 = no, and 1226 = NA.
 Notably, the GSS codes missing values with multiple indicators--".i: Inapplicable," ".n: No answer," ".d: Do not Know/Cannot Choose," and ".s: Skipped on Web." For the purposes of our preliminary study, we coded all missing data as NA.
@@
 
-Our control and independent variables are representative of two different dimensions we sought to capture. The first dimension is individual-level or identity characteristics:
+Our control and independent variables are representative of two different dimensions we sought to capture. The first dimension is individual-level or social identity characteristics:
 
@@
 
--   **age**
+-   **age**: Respondent's age in numeric values.
 
 -   **sexnow1**: Do you describe yourself as male, female, or transgender?
-
--   **sex**: Code respondent's sex
 
@@
 Similarly, gender was a contentious variable to account for.
-Although the GSS has several questions aimed at capturing respondent's gender, we found them to be lacking in breathe.
+Although the GSS has several questions aimed at capturing respondent's gender, we found them to be lacking in breadth.
 For example, the GSS codes **sex** within a Male/Female binary based solely on interviewer discretion.
@@
 Out of 2306 cases, only 10 were coded as Transgender and 9 as None of these.
+Therefore, this is by no means a representative sample on gender minority population.
+The fact that the GSS data is primarily collected through in-person field interviews increases the possibility that minority-identifying individuals may not disclose their accurate identity information to avoid being associated with negative stereotypes.
 
@@
 Out of 2318 cases, 2164 were coded as NAs.
+In future work, we will pursue data options that allow the analysis ton include citizenship status.
 
-The second dimension we sought to capture was community-level characteristics, i.e., the social location of individuals. For this dimension, we drew on the following variables:
+The second dimension we sought to capture was community-level characteristics that would potentially affect one's sense of safety. For this dimension, we drew on the following variables:
 
@@
 -   **raclive**: Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?
+
+These variables seek to account for respondent's social embeddedness and networks of support.
+Specifically, **raclive** is an important variable to consider as it accounts for neighborhood composition, in the absence of a neighborhood dissimilarity index.
+While this variable only looks at Black-white racial composition, it provides a good benchmark to examine broader racial antagonisms.
+Research has documented the associated fears white experience when Blacks or people of other racial backgrounds move into their neighborhood [@lacayo2017; @massey2020].
+
+The third dimension we included were on :
 
@@
 -   **partnrs5**: since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?
-
-These variables seek to account for respondent's social embeddedness and networks of support.
-Specifically, **raclive** is an important variable to consider as it accounts for neighborhood composition, in the absence of a neighborhood dissimilarity index.
-While this variable only looks at Black-white racial composition, it provides a good benchmark to examine broader racial antagonisms.
-Research has documented the associated fears white experience when Blacks or people of other racial backgrounds move into their neighborhood [@lacayo2017; @massey2020].
 
@@
 ```
+
 @fig-FearRaceGender does a good job at plotting this relationship.
@@
 ```
+
 # Discussion
@@
 \newpage
+
 # References

### 2023-12-06 scripts/manuscript.qmd
recategorizing variables of interests

@@
 
-The third dimension we included were on :
+The third dimension we considered were individual-level health-related and  attitudinal variables.
+Good mental and physical health status is positively correlated with individuals' evaluation of their life conditions and affects how likely individuals are to report positive emotions about their life [@dwyer2017].
+Additionally, social capital--such as levels of interpersonal trust and norms of reciprocity and mutual aid--is believed to play an important role in explaining variations in health status across geographic localities [@kawachi1999].
+Therefore, we included the following variables that capture respondents' self-reported health status and assessment of interpersonal social climate:
 

### 2023-12-06 scripts/manuscript.qmd
some fear of crime lit

@@
 format: html
+    wordcount-html: default
 toc: true
@@
 
-I'm leaving myself a note here to figure out how to do word count in a Quarto document.
-This will allow me to directly write in the qmd file and save the trouble of writing elsewhere like a word document and pasting things here.
+This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue.
+With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
+
+Individuals' report of fear have been shown to relate to a wide range of contextual factors, such as neighborhood composition and community crime rates [@hale1996; @warr1990], and individual sociodemographic characteristics, such as gender, sexual orientation, income, age, health condition, and immigration status [@lagrange1989; @otis2007; @madriz2023].
+
+Especially, the role of gender in fear of crime has been well documented in victimization research [@stanko1993; @snedker2006; @schnitzer2010].
 

### 2023-12-06 scripts/manuscript.qmd
contextual factors lit

@@
 Individuals' report of fear have been shown to relate to a wide range of contextual factors, such as neighborhood composition and community crime rates [@hale1996; @warr1990], and individual sociodemographic characteristics, such as gender, sexual orientation, income, age, health condition, and immigration status [@lagrange1989; @otis2007; @madriz2023].
+To be more specific, this body of scholarship has found spatial and neighborhood characteristics have a significant impact on individuals' self-reported fear of crime, as those who live in areas with high levels of community disorder have reported more fear than those who do not live in such communities [@rountree1996].
+Also, due to the popularity density in urban areas, residents of these areas have reported higher levels of fear at night, potentially because of feelings of alienation and unfamiliarity among residents [@bennett1994].
 
-Especially, the role of gender in fear of crime has been well documented in victimization research [@stanko1993; @snedker2006; @schnitzer2010].
+Especially, the role of gender in fear of crime has been well documented in victimization research [@cops2011; @snedker2006; @schnitzer2010].
+
+[@stanko1993]
 

### 2023-12-06 scripts/manuscript.qmd
gender variable

@@
 
-Especially, the role of gender in fear of crime has been well documented in victimization research [@cops2011; @snedker2006; @schnitzer2010].
-
-[@stanko1993]
+Especially, the role of gender in fear of crime has been well documented in victimization research [@cops2011; @snedker2006; @schnitzer2010; @stanko1993].
+Early scholarship emphasized that women tend to report more fear than men  [@stanko1990], with some initial research investigating the seeming disjuncture between women's fear and their actual experiences of victimization, as women were often found to report high levels of fear for crimes that they were at relatively little risk of experiencing [@hough1983].
+Subsequent studies suggested that women's fear should be understood in the context of the many potential threats that women confront, including their heightened risk of sexual assault [@stanko1993].
+Therefore, fear at night is gradually perceived as a manifestation of gender-based violence with sexual assault implications, though acknowledging that sexual assault is merely one among many potential forms of crime.
 

### 2023-12-06 scripts/manuscript.qmd
background

@@
 
-This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue.
-With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
-
 Individuals' report of fear have been shown to relate to a wide range of contextual factors, such as neighborhood composition and community crime rates [@hale1996; @warr1990], and individual sociodemographic characteristics, such as gender, sexual orientation, income, age, health condition, and immigration status [@lagrange1989; @otis2007; @madriz2023].
@@
 Therefore, fear at night is gradually perceived as a manifestation of gender-based violence with sexual assault implications, though acknowledging that sexual assault is merely one among many potential forms of crime.
+However, the more insidious and everyday forms of harassment are frequently not captured by common survey instruments--they regularly privilege the more obvious or extreme forms of victimization and fail to account for the routine forms of violence that women confront [@moran2003].
+We focus our attention on this less intrusive but omnipresent form of everyday violence and examine the unequal effects it has on different populations.
+We ask: what groups of people are more likely to have experienced fear of walking alone at night in their neighborhood?
+
+Historically, despite all the aforementioned contexts and individual-level characteristics being identified as salient to individuals’ self-reported fear at night, they were usually studied in isolation.
+Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously.
+Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals's sense of threat from social surroundings, i.e., fear at night.
+This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue.
+With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
 

### 2023-12-06 scripts/manuscript.qmd
Intro

@@
 Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously.
-Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals's sense of threat from social surroundings, i.e., fear at night.
+Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.
+
 This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue.
 With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
+We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walkign alone at night" and our other variables of interest.
 

### 2023-12-06 scripts/manuscript.qmd
first half

@@
 Therefore, fear at night is gradually perceived as a manifestation of gender-based violence with sexual assault implications, though acknowledging that sexual assault is merely one among many potential forms of crime.
-However, the more insidious and everyday forms of harassment are frequently not captured by common survey instruments--they regularly privilege the more obvious or extreme forms of victimization and fail to account for the routine forms of violence that women confront [@moran2003].
+However, the more insidious and everyday forms of harassment are frequently not captured by common survey instruments----they regularly privilege the more obvious or extreme forms of victimization and fail to account for the routine forms of violence that women confront [@moran2003].
 We focus our attention on this less intrusive but omnipresent form of everyday violence and examine the unequal effects it has on different populations.
@@
 
-Historically, despite all the aforementioned contexts and individual-level characteristics being identified as salient to individuals’ self-reported fear at night, they were usually studied in isolation.
+Historically, despite all the aforementioned contexts and individual-level characteristics being identified as salient to individuals' self-reported fear at night, they were usually studied in isolation.
 Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously.
-Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.
+To fill this gap in knowledge, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear----intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.
 
 This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue.
-With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
+With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among the US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
 We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walkign alone at night" and our other variables of interest.
@@
 
--   **sexornt**: Which of the following best describes you?
+-   **sexornt**: Which of the following best describes you? [1] "Gay, lesbian, or homosexual" [2] Bisexual [3] Heterosexual or straight [99] "Refused"
 
@@
 
--   **degree**: Respondent's degree.
+-   **degree**: Respondent's (highest educational) degree.
 
@@
 Good mental and physical health status is positively correlated with individuals' evaluation of their life conditions and affects how likely individuals are to report positive emotions about their life [@dwyer2017].
-Additionally, social capital--such as levels of interpersonal trust and norms of reciprocity and mutual aid--is believed to play an important role in explaining variations in health status across geographic localities [@kawachi1999].
+Additionally, social capital----such as levels of interpersonal trust and norms of reciprocity and mutual aid--is believed to play an important role in explaining variations in health status across geographic localities [@kawachi1999].
 Therefore, we included the following variables that capture respondents' self-reported health status and assessment of interpersonal social climate:

### 2023-12-06 scripts/manuscript.qmd
inline code requirement fulfilled

@@
 
--   **sexornt**: Which of the following best describes you? [1] "Gay, lesbian, or homosexual" [2] Bisexual [3] Heterosexual or straight [99] "Refused"
+-   **sexornt**: Which of the following best describes you? \[1\] "Gay, lesbian, or homosexual" \[2\] Bisexual \[3\] Heterosexual or straight \[99\] "Refused"
 
@@
 We seek to show how the experience of fear differed across different measures of individual-level and community-level characteristics.
-To begin with, across our sample of 2318 cases, mean fear was calculated at `mean_fear`.
+To begin with, across our sample of 2318 cases, mean fear was calculated at `r mean(as.numeric(overall_fear$fear, na.rm=TRUE))`.
 

### 2023-12-06 scripts/manuscript.qmd
results 1

@@
 To begin with, across our sample of 2318 cases, mean fear was calculated at `r mean(as.numeric(overall_fear$fear, na.rm=TRUE))`.
+One thing to note about the interpretation of this value is, because we operationalized the variable **fear** as a dichotomous variable with answers being "Yes = 1" and "No = 0", this mean value approximates "the likelihood for an individual in our sample to report the experience of fear of walking alone at night in their neighborhood".
 
 First, we examined how fear is experienced across race and gender.
-As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents that identified as Male.
+As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents who identified as Male.
 Particularly, respondents that were Transgender expressed greater fear, but this may be indicative of sample representation.
@@
 When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.
+This finding is consistent with the hypothesis that the most privileged social group in terms of race has the lowest level of fear.
 

### 2023-12-06 scripts/manuscript.qmd
transgender

@@
 With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among the US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
-We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walkign alone at night" and our other variables of interest.
+We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walking alone at night" and our other variables of interest.
 
@@
 For example, only one respondent identified as both Hispanic and Transgender, and since they responded affirmative to the question, this finding is heavily skewed.
+However, it's important to note that Transgender individuals have been subject to more frequent and more severe forms of sexual violence and thus, experiencing more fear related to gender-based crime and engaging in more gendered tactics to protect themselves [@schilt2009].
+
 When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.

### 2023-12-06 scripts/manuscript.qmd
Results from Table 1 and Figure 1

@@
 
-When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.
-This finding is consistent with the hypothesis that the most privileged social group in terms of race has the lowest level of fear.
-
 ```{r}
@@
 ```
+
+When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.
+This finding is aligned with the hypothesis that the most privileged social group in terms of race has the lowest level of fear.
+Additionally, our visualizations show the compounding effects of both race and gender----consistent with the framework of intersectionality [@mccall2005]----that the rates of reporting fear at night is higher when the group is marginalized both racially and gender-wise.
+For example, except for the Transgender respondents, NHPI Females, Asian Females and AIAN females reported the highest rates of fear, while White males reported the lowest rates of fear.
 

### 2023-12-06 scripts/manuscript.qmd
social embeddedness

@@
 Moving on to measures of social embeddedness, we examined fear across degrees of religious participation, degree of happiness and health, residential diversity, and sexual activity.
-Regarding religious participation, we did not discern any clear pattern, as @fig-FearReligiousPart displays below.
+Regarding religious participation, we did not discern any clear pattern.
+Across all frequencies of religious activity attendance, ranging from "Never" to "Nearly every week", the percentages of respondents who reported fear at night were all close to 40%, as @fig-FearReligiousPart displays below.
+The one outlier is attending religious services "about once a month", which we have no theoretical backings for and is likely attributable to sampling issues.
 

### 2023-12-06 scripts/manuscript.qmd
residential composition

@@
 
-Next, looking at measures of Happiness and Health, we found that generally as happiness and health increase, social fear decreases. This is shown in @fig-FearHappiness and @fig-FearHealth.
+Next, looking at residential composition and race combined displayed in @fig-FearNeighborhood, we do notice a pattern where respondents living in residentially more diverse areas were more likely to respond affirmative to the question of social fear.
+This indicates the sentiment that respondents felt safer and more comfortable living in neighborhoods where other residents share their racial identities, which speaks to
+
+```{r}
+#| label: fig-FearNeighborhood
+#| fig-cap: Bar graph showing the percentage of respondants, based on race and neighborhood diversity, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
+
+gss_subset %>% 
+  filter(!is.na(racecombo)) %>% 
+  filter(racecombo != "Other") %>% 
+  filter(!is.na(raclive)) %>%  
+  group_by(racecombo, raclive) %>% 
+  summarize(fear = sum(fear == "yes"), 
+            n = n(),
+            prop = fear/n) %>% 
+  ggplot(aes(x = racecombo, y = prop)) +
+  geom_col(aes(fill = raclive), position = "dodge") + 
+  scale_fill_brewer(palette = "Set2") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Social Fear by Race and Neighborhood Diversity",
+       subtitle = "Is social fear influenced by respondent's race and their neighborhood's diversity?",
+       x = "Race",
+       y = "Percentage of Respondents who Expressed Fear") +
+  theme_bw() +
+  guides(fill = guide_legend(title = "Residential Diversity")) 
+```
+
+What's more, looking at measures of individual-level characteristics like Happiness and Health, we found that generally as happiness and health increase, social fear decreases.
+This is shown in @fig-FearHappiness and @fig-FearHealth.
+This finding resonates with recent scholarship on the direct relationship between fear of crime and subject well-being that higher levels of happiness and better physical health are associated with lower levels of fear [@krulichova2018].
 
@@
 
-Next, looking at residential composition and race combined displayed in @fig-FearNeighborhood, we do notice a pattern where respondents living in residentially more diverse areas were more likely to respond affirmative to the question of social fear.
-
-```{r}
-#| label: fig-FearNeighborhood
-#| fig-cap: Bar graph showing the percentage of respondants, based on race and neighborhood diversity, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?
-
-gss_subset %>% 
-  filter(!is.na(racecombo)) %>% 
-  filter(racecombo != "Other") %>% 
-  filter(!is.na(raclive)) %>%  
-  group_by(racecombo, raclive) %>% 
-  summarize(fear = sum(fear == "yes"), 
-            n = n(),
-            prop = fear/n) %>% 
-  ggplot(aes(x = racecombo, y = prop)) +
-  geom_col(aes(fill = raclive), position = "dodge") + 
-  scale_fill_brewer(palette = "Set2") +
-  scale_y_continuous(labels = scales::percent) +
-  labs(title = "Social Fear by Race and Neighborhood Diversity",
-       subtitle = "Is social fear influenced by respondent's race and their neighborhood's diversity?",
-       x = "Race",
-       y = "Percentage of Respondents who Expressed Fear") +
-  theme_bw() +
-  guides(fill = guide_legend(title = "Residential Diversity")) 
-```
-
 Lastly, we looked at sexual activity as a marker of social fear as represented in @fig-FearPartnersGender.

### 2023-12-07 scripts/manuscript.qmd
Discussion

@@
 Next, looking at residential composition and race combined displayed in @fig-FearNeighborhood, we do notice a pattern where respondents living in residentially more diverse areas were more likely to respond affirmative to the question of social fear.
-This indicates the sentiment that respondents felt safer and more comfortable living in neighborhoods where other residents share their racial identities, which speaks to
+This indicates the sentiment that respondents felt safer and more comfortable living in neighborhoods where other residents share their racial identities, which speaks to the reality of racial residential segregation and its impact on health [@charles2003; @williams2016].
 
@@
 Lastly, we looked at sexual activity as a marker of social fear as represented in @fig-FearPartnersGender.
+Overall, it could be implied that having more sexual partners is positively associated with experiencing fear of walking alone at night, especially for indivuals who reported having had over 20 sexual partners in the last 5 years.
+A potential explanation is that having more than 20 sexual partners is indicative of a more sexually exploratory and adventurous lifestyle, which brings more exposure to unfamiliar social contacts and environments.
+We take notice that, interestingly, Females with less than 5 sexual partners in the last 5 years reported lower rates of fear at night compared to Females who reported having 0 sexual partneres.
+Our interpretation is that those in the "Less than 5 (sexual partners)" group are more likely to have had stable, longer-term sexual relationships, which may have brought a certain sense of security and even cohabiting partners.
 

### 2023-12-07 scripts/manuscript.qmd
complete

@@
   - W. Jamie Yang
-abstract: "TBW" # this is where abstract should be pasted 
+abstract: "Fear at night has been perceived as a form of gender-based violence with sexual assault implications.
+However, this more insidious and everyday forms of harassment are frequently not captured by common survey instruments.
+In this study, we investigate the inequalities in terms of fear of walking alone at night among the US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.
+We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between fear of walking alone at night and our other variables of interest.
+Our findings suggest that fear at night differs by race, gender, subjective health, and neighborhood diversity.
+Consistent with previous research, women were also more likely than men to report fear at night.
+Importantly, groups with multiply marginalized social identities are likely to report higher rates of fear at night." # this is where abstract should be pasted
 format: html
@@
 execute:
-  eval: true # change to false before final submission
+  eval: false # change to false before final submission
   echo: true
@@
 Therefore, fear at night is gradually perceived as a manifestation of gender-based violence with sexual assault implications, though acknowledging that sexual assault is merely one among many potential forms of crime.
-However, the more insidious and everyday forms of harassment are frequently not captured by common survey instruments----they regularly privilege the more obvious or extreme forms of victimization and fail to account for the routine forms of violence that women confront [@moran2003].
+However, the more insidious and everyday forms of harassment are frequently not captured by common survey instruments---they regularly privilege the more obvious or extreme forms of victimization and fail to account for the routine forms of violence that women confront [@moran2003].
 We focus our attention on this less intrusive but omnipresent form of everyday violence and examine the unequal effects it has on different populations.
@@
 Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously.
-To fill this gap in knowledge, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear----intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.
+To fill this gap in knowledge, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear---intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.
 
@@
 
+Previous research on fear at night has attended to differences between women and men, and among sexual minority populations [@meyer2014; otis2007].
+This line of research, however, has not fully examined the intersectional effects of various social dimensions on fear at night, such as race, sexual orientation, gender, and community composition.
+To be specific, although significant gender differences have been documented [@cops2011; @stanko1990], wherein women are more likely than men to report fear, research has failed to consider the complex ways in which other social dimensions such as race may intersect with regard to such fear.
+
+In this project, we used nationally representative survey data to examine racial, sexual, gender, health, and community contextual differences in reports of being afraid to walk alone at night.
+Our findings suggest that fear at night differs by race, gender, subjective health, and neighborhood diversity.
+Consistent with previous research, women were also more likely than men to report fear at night.
+Further analyses revealed that there exists a significant interaction between race and gender in report of fear.
+Importantly, groups with multiply marginalized social identities are likely to report higher rates of fear at night.
+
+Our assessment of fear relies on a single measure---fear of walking alone at night---that does not explicitly ask respondents about what, in particular, they fear.
+Potential objects or events in question would be robbery, sexual assault, verbal harassment, or sexual orientation- or religion-based hate crimes.
+And with common sense, we know that people embodying different identities and self-presentations will be vulnerable to different forms of street crimes.
+Further research should explore the particular nature of this fear, and whether it is based on crime victimization experience or verbal harassment.
+
+Despite limitations of this study, our findings provide a good understanding of sociodemographic differences in fear at night.
+We uncovered an important relation between fear and disadvantaged status: racial and ethnic minorities, women, and individuals of better subjective health are more likely to report fear of victimization than their privileged counterparts.
+
+The R tools we learned from the class "EDLD 651: Intro to R" have been extremely helpful in cleaning and visualizing the data.
+
 \newpage

### 2023-12-07 scripts/manuscript.qmd
try commit final draft

@@
 execute:
-  eval: false # change to false before final submission
+  eval: true # change to false before final submission
   echo: true

### 2023-12-07 scripts/manuscript.qmd
edit

@@
 Consistent with previous research, women were also more likely than men to report fear at night.
-Importantly, groups with multiply marginalized social identities are likely to report higher rates of fear at night." # this is where abstract should be pasted
+Importantly, groups with multiply marginalized social identities are likely to report higher rates of fear at night."
 format: html
@@
 execute:
-  eval: true # change to false before final submission
+  eval: true # if we change to false, in-line code will not run 
   echo: true
